---
title: "Lab Exp (RS reviewed)"
format: html
editor: source
---

```{r}
rm(list = ls())
pacman::p_load("tidyverse")
```

```{r}
dat <- read.csv("behavior-18.csv")
dat_demo <- read.csv("demographics_2.csv")
```


# demografic
```{r}
# one of them is test data (age = 100)
exclude_ids <- dat_demo |>
  filter(age >= 100) |>
  pull(user_id)

dat <- dat |>
  filter(!(gsub("-", "", user_id) %in% exclude_ids))

# gender
dat_demo <- dat_demo |>
  filter(age < 100)

table(dat_demo$sex)

# age
mean(dat_demo$age)
sd(dat_demo$age)
table(dat_demo$age)

dplyr::n_distinct(dat$user_id)
```


# excluding data
```{r}
# trial
dat_analysis <- dat |>
  mutate(round = round + 1)
summary(dat_analysis$round)

# check
stopifnot(all(table(dat_analysis$user_id) == 13))

# except practice trials
dat_analysis <- dat_analysis |>
  filter(round >= 4)


glimpse(dat_analysis)
```


```{r}
boxplot(dat_analysis$explore_count) # OK

# emotional rating check
straightline_rows <- dat_analysis |>
  dplyr::filter(
    satisficing == disappointment,
    disappointment == successful,
    successful == outer.regret,
    outer.regret == inner.regret
  )
nrow(straightline_rows)

table(dat_analysis$explore_count)

# storage error
dat_analysis |>
  dplyr::filter(explore_count == 0) |>
  dplyr::select(user_id, round, explore_count, explore_rt, survey_rt)

dat_analysis <- dat_analysis |>
  dplyr::filter(explore_count >= 1) # storage errors

# explore_count_next / explore_count_diff are pre-computed in the raw CSV,
# so the trial just before a storage error still carries k_{t+1} = 0
dat_analysis |>
  dplyr::filter(explore_count_next == 0) |>
  dplyr::select(
    user_id, round, explore_count,
    explore_count_next, explore_count_diff
  )

dat_analysis <- dat_analysis |>
  dplyr::mutate(
    explore_count_diff = replace(
      explore_count_diff,
      explore_count_next %in% 0, NA
    ),
    explore_count_next = replace(
      explore_count_next,
      explore_count_next %in% 0, NA
    )
  )


boxplot(dat_analysis$explore_count) # OK
glimpse(dat_analysis)
```


# Subjective regret reflects objective breadth and depth deficits

## Objective loss and subjective regret
```{r}
# objective loss
library(dplyr)

dat_analysis <- dat_analysis |>
  mutate(
    # depth-deficit loss
    theo_inner_regret = observed_best - user_chosen_point,

    # breadth-deficit loss
    theo_outer_regret = global_best - observed_best
  )

glimpse(dat_analysis)
summary(dat_analysis$theo_inner_regret)
summary(dat_analysis$theo_outer_regret)
```


## Between losses and subjective emotions
```{r}
library(lme4)
library(lmerTest)

dat_analysis <- dat_analysis |>
  mutate(
    z_theo_outer_regret = as.numeric(scale(theo_outer_regret)),
    z_theo_inner_regret = as.numeric(scale(theo_inner_regret)),
    z_outer_regret = as.numeric(scale(outer.regret)),
    z_inner_regret = as.numeric(scale(inner.regret)),
    z_disappointment = as.numeric(scale(disappointment)),
    z_satisfaction = as.numeric(scale(satisficing)),
    z_successful = as.numeric(scale(successful)),
    z_round = as.numeric(scale(round))
  )

# breadth-deficit loss and subjective breadth regret
breadthloss_breadthsubject <- lmer(
  z_outer_regret ~ z_theo_outer_regret + (z_theo_outer_regret | user_id),
  data = dat_analysis
)

summary(breadthloss_breadthsubject)

# breadth-deficit loss and subjective depth regret
breadthloss_depthsubject <- lmer(
  z_inner_regret ~ z_theo_outer_regret + (z_theo_outer_regret | user_id),
  data = dat_analysis
)

summary(breadthloss_depthsubject)


# depth-deficit loss and subjective depth regret
depthloss_depthsubject <- lmer(
  z_inner_regret ~ z_theo_inner_regret + (z_theo_inner_regret | user_id),
  data = dat_analysis
)

summary(depthloss_depthsubject)

# depth-deficit loss and subjective breadth regret
depthloss_breadthsubject <- lmer(
  z_outer_regret ~ z_theo_inner_regret + (z_theo_inner_regret | user_id),
  data = dat_analysis
)

summary(depthloss_breadthsubject)
```

```{r}
library(ggplot2)
library(patchwork)

slope_pvalue <- function(model) {
  co <- summary(model)$coefficients
  if ("Pr(>|t|)" %in% colnames(co)) {
    co[2, "Pr(>|t|)"] # lmerTest::lmer
  } else if ("p-value" %in% colnames(co)) {
    co[2, "p-value"] # nlme::lme
  } else {
    NA_real_
  }
}

stars <- function(p) {
  if (is.na(p)) {
    ""
  } else if (p < .001) {
    "***"
  } else if (p < .01) {
    "**"
  } else if (p < .05) {
    "*"
  } else {
    " (n.s.)"
  }
}

plot_lmm <- function(
  model,
  x_var,
  y_var,
  data,
  title,
  significant = NULL,
  x_label = x_var,
  y_label = y_var,
  base_size = 11
) {
  indiv_coefs <- as.data.frame(coef(model)$user_id)
  names(indiv_coefs) <- c("b0", "b1")
  fe <- fixef(model)

  p_val <- slope_pvalue(model)
  if (is.null(significant)) {
    significant <- !is.na(p_val) && p_val < .05
  }

  line_col <- if (significant) "#2166ac" else "#aaaaaa"
  fe_col <- if (significant) "#d73027" else "#555555"
  subtitle <- paste0("β = ", round(fe[2], 2), stars(p_val))

  ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(alpha = 0.15, size = 0.7, color = "grey40") +
    geom_abline(
      data = indiv_coefs,
      aes(intercept = b0, slope = b1),
      alpha = 0.25,
      color = line_col,
      linewidth = 0.3
    ) +
    geom_abline(
      intercept = fe["(Intercept)"],
      slope = fe[2],
      color = fe_col,
      linewidth = 1.2
    ) +
    labs(title = title, subtitle = subtitle, x = x_label, y = y_label) +
    theme_classic(base_size = base_size) +
    theme(
      plot.title = element_text(size = base_size - 1, face = "bold"),
      plot.subtitle = element_text(size = base_size - 2)
    )
}
```

```{r}
# each loss and subjective regret
p1 <- plot_lmm(
  breadthloss_breadthsubject,
  "z_theo_outer_regret",
  "z_outer_regret",
  dat_analysis,
  "Breadth loss → Breadth regret",
  TRUE,
  x_label = "Breadth-deficit loss (z)",
  y_label = "Breadth regret (z)"
)

p2 <- plot_lmm(
  breadthloss_depthsubject,
  "z_theo_outer_regret",
  "z_inner_regret",
  dat_analysis,
  "Breadth loss → Depth regret",
  FALSE,
  x_label = "Breadth-deficit loss (z)",
  y_label = "Depth regret (z)"
)

p3 <- plot_lmm(
  depthloss_breadthsubject,
  "z_theo_inner_regret",
  "z_outer_regret",
  dat_analysis,
  "Depth loss → Breadth regret",
  FALSE,
  x_label = "Depth-deficit loss (z)",
  y_label = "Breadth regret (z)"
)

p4 <- plot_lmm(
  depthloss_depthsubject,
  "z_theo_inner_regret",
  "z_inner_regret",
  dat_analysis,
  "Depth loss → Depth regret",
  TRUE,
  x_label = "Depth-deficit loss (z)",
  y_label = "Depth regret (z)"
)

(p1 | p2) / (p3 | p4)
```

```{r}
# sum of subjective regret and gain point
dat_analysis <- dat_analysis |>
  mutate(
    emo_sum = outer.regret + inner.regret,
    z_emo_sum = as.vector(scale(emo_sum)),
    z_point = as.vector(scale(user_chosen_point))
  )

# regret sum → gain point
regretsum_point <- lmer(
  z_point ~ z_emo_sum + (z_emo_sum | user_id),
  data = dat_analysis
)
summary(regretsum_point)

plot_lmm(
  regretsum_point,
  "z_emo_sum",
  "z_point",
  dat_analysis,
  "Sum of regret → Gain point",
  x_label = "Sum of subjective regret (z)",
  y_label = "Gain point (z)",
  base_size = 14
)
```



### SI: other emotions

```{r}
fit_emotion_loss <- function(emotion, loss) {
  f <- reformulate(
    c(loss, sprintf("(%s | user_id)", loss)),
    response = emotion
  )
  lmer(f, data = dat_analysis)
}

emotions <- c("z_disappointment", "z_satisfaction", "z_successful")
losses <- c("z_theo_inner_regret", "z_theo_outer_regret")

si_emotion_models <- tidyr::expand_grid(emotion = emotions, loss = losses) |>
  purrr::pmap(\(emotion, loss) {
    m <- fit_emotion_loss(emotion, loss)
    co <- summary(m)$coefficients
    tibble::tibble(
      emotion = emotion,
      loss = loss,
      beta = co[2, "Estimate"],
      SE = co[2, "Std. Error"],
      df = co[2, "df"],
      t = co[2, "t value"],
      p = co[2, "Pr(>|t|)"],
      singular = lme4::isSingular(m)
    )
  }) |>
  purrr::list_rbind()

si_emotion_models
```


### The Relationship Between the Two Types of Regret

```{r}
dat_analysis <- dat_analysis |>
  mutate(
    theo_diff = theo_inner_regret - theo_outer_regret,
    emo_diff = inner.regret - outer.regret
  )

ggplot(dat_analysis, aes(x = theo_diff)) +
  geom_point(
    aes(y = inner.regret),
    alpha = 0.4,
    size = 2,
    color = "skyblue"
  ) +
  geom_smooth(
    aes(y = inner.regret),
    method = "loess",
    se = FALSE,
    color = "skyblue",
    linetype = "dashed"
  ) +
  geom_point(
    aes(y = outer.regret),
    alpha = 0.4,
    size = 2,
    color = "orange"
  ) +
  geom_smooth(
    aes(y = outer.regret),
    method = "loess",
    se = FALSE,
    color = "orange",
    linetype = "dashed"
  ) +
  geom_vline(xintercept = 0) +
  labs(
    x = "Theoretical regret difference (inner − outer)",
    y = "Subjective regret",
    title = "Correspondence between theoretical and subjective regret"
  ) +
  theme_minimal(base_size = 14)
```


# Sequential adjustment driven by regret signals

```{r}
# change in search set size ~ regret
library(nlme)
library(car)

fixed_formula <- explore_count_diff ~ z_inner_regret + z_outer_regret + z_round
ctrl <- lmeControl(opt = "optim", maxIter = 200)

lm_null <- gls(
  fixed_formula,
  data = dat_analysis,
  method = "ML",
  na.action = na.omit
)
lmm_intercept <- lme(
  fixed_formula,
  random = ~ 1 | user_id,
  data = dat_analysis,
  method = "ML",
  na.action = na.omit
)
lmm_slopes_diag <- lme(
  fixed_formula,
  random = list(user_id = pdDiag(~ z_inner_regret + z_outer_regret)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
lmm_slopes <- lme(
  fixed_formula,
  random = ~ z_inner_regret + z_outer_regret | user_id,
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)

anova(lm_null, lmm_intercept, lmm_slopes_diag, lmm_slopes)

summary(lmm_slopes_diag)
Anova(lmm_slopes_diag, type = "3")
```

## β-visualization

```{r}
summary(lmm_slopes_diag)$tTable |>
  as.data.frame() |>
  tibble::rownames_to_column("term") |>
  filter(term %in% c("z_outer_regret", "z_inner_regret")) |>
  mutate(
    regret_type = ifelse(
      term == "z_outer_regret",
      "Subjective breadth regret",
      "Subjective depth regret"
    ),
    ci_low = Value + qt(0.025, df = 500) * Std.Error,
    ci_high = Value + qt(0.975, df = 500) * Std.Error
  ) |>
  ggplot(aes(x = regret_type, y = Value, fill = regret_type)) +
  geom_col(width = 0.6, alpha = 0.85) +
  geom_jitter(
    data = ranef(lmm_slopes_diag) |>
      mutate(
        `Subjective breadth regret` = z_outer_regret +
          fixef(lmm_slopes_diag)["z_outer_regret"],
        `Subjective depth regret` = z_inner_regret +
          fixef(lmm_slopes_diag)["z_inner_regret"]
      ) |>
      pivot_longer(
        c(`Subjective breadth regret`, `Subjective depth regret`),
        names_to = "regret_type",
        values_to = "slope"
      ),
    aes(x = regret_type, y = slope),
    width = 0.15,
    alpha = 0.4,
    size = 1.8,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.15) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_fill_manual(
    values = c(
      "Subjective breadth regret" = "darkorange3",
      "Subjective depth regret" = "steelblue3"
    )
  ) +
  labs(x = NULL, y = expression(beta ~ "predicting" ~ Delta * k)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
```

## Other emotion
```{r}
lmm_satisfaction <- lme(
  fixed = explore_count_diff ~
    z_satisfaction +
    z_round,
  random = ~ 1 | user_id,
  data = dat_analysis,
  method = "ML",
  na.action = na.omit
)
summary(lmm_satisfaction)

Anova(lmm_satisfaction, type = "3")


lmm_disappoint <- lme(
  fixed = explore_count_diff ~
    z_disappointment +
    z_round,
  random = ~ 1 | user_id,
  data = dat_analysis,
  method = "ML",
  na.action = na.omit
)
summary(lmm_disappoint)

Anova(lmm_disappoint, type = "3")


lmm_success <- lme(
  fixed = explore_count_diff ~
    z_successful +
    z_round,
  random = ~ 1 | user_id,
  data = dat_analysis,
  method = "ML",
  na.action = na.omit
)
summary(lmm_success)

Anova(lmm_success, type = "3")
```

## Other emotion-randam slope
```{r}
lmm_satisfaction_si <- lme(
  explore_count_diff ~ z_satisfaction + z_round,
  random = list(user_id = pdDiag(~z_satisfaction)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
summary(lmm_satisfaction_si)


lmm_disappoint_si <- lme(
  explore_count_diff ~ z_disappointment + z_round,
  random = list(user_id = pdDiag(~z_disappointment)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
summary(lmm_disappoint_si)


lmm_success_si <- lme(
  explore_count_diff ~ z_successful + z_round,
  random = list(user_id = pdDiag(~z_successful)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
summary(lmm_success_si)
```

## Robustness check: controlling for current search-set size (k_t)

```{r}
dat_analysis <- dat_analysis |>
  mutate(
    z_k = as.numeric(scale(explore_count)),
    z_k2 = as.numeric(scale(explore_count^2))
  )

lmm_slopes_diag_kt <- lme(
  explore_count_diff ~ z_k + z_k2 + z_inner_regret + z_outer_regret + z_round,
  random = list(user_id = pdDiag(~ z_inner_regret + z_outer_regret)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
summary(lmm_slopes_diag_kt)
Anova(lmm_slopes_diag_kt, type = "3")
```


### Exploratory: additionally controlling for objective losses

```{r}
# breadth-/depth-deficit loss entered alongside both regrets and k_t (+ k_t^2)
lmm_slopes_diag_kt_loss <- lme(
  explore_count_diff ~ z_k +
    z_k2 +
    z_inner_regret +
    z_outer_regret +
    z_theo_inner_regret +
    z_theo_outer_regret +
    z_round,
  random = list(user_id = pdDiag(~ z_inner_regret + z_outer_regret)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
summary(lmm_slopes_diag_kt_loss)
Anova(lmm_slopes_diag_kt_loss, type = "3")
```


### Same k_t control applied to the other subjective responses (specificity check)

```{r}
lmm_satisfaction_si_kt <- lme(
  explore_count_diff ~ z_k + z_k2 + z_satisfaction + z_round,
  random = list(user_id = pdDiag(~z_satisfaction)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
summary(lmm_satisfaction_si_kt)
Anova(lmm_satisfaction_si_kt, type = "3")


lmm_disappoint_si_kt <- lme(
  explore_count_diff ~ z_k + z_k2 + z_disappointment + z_round,
  random = list(user_id = pdDiag(~z_disappointment)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
summary(lmm_disappoint_si_kt)
Anova(lmm_disappoint_si_kt, type = "3")


lmm_success_si_kt <- lme(
  explore_count_diff ~ z_k + z_k2 + z_successful + z_round,
  random = list(user_id = pdDiag(~z_successful)),
  data = dat_analysis,
  method = "ML",
  control = ctrl,
  na.action = na.omit
)
summary(lmm_success_si_kt)
Anova(lmm_success_si_kt, type = "3")
```


# Behavioral adaptation toward the theoretical optimum

### Initial classification (Above/Below optimum)
```{r}
initial_k <- dat_analysis |>
  group_by(user_id) |>
  slice_min(round, n = 1, with_ties = FALSE) |>
  ungroup() |>
  dplyr::select(user_id, initial_round = round, initial_k = explore_count)

initial_k |> count(initial_round)

dat_analysis <- dat_analysis |>
  left_join(initial_k, by = "user_id") |>
  mutate(
    group = case_when(
      is.na(initial_k) ~ NA_character_,
      initial_round != 4 ~ NA_character_,
      initial_k < 16 ~ "Below optimum", # simulation-derived benchmark k* = 16
      TRUE ~ "Above optimum"
    )
  )

dat_analysis |>
  distinct(user_id, group, initial_k) |>
  count(group)
```

### Visualization and interaction
```{r}
dat_analysis |>
  filter(!is.na(group), group != "At optimum") |>
  group_by(group, round) |>
  summarise(
    mean_k = mean(explore_count, na.rm = TRUE),
    se_k = sd(explore_count, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
  mutate(trial = round - 3) |>
  ggplot(aes(x = trial, y = mean_k, color = group, fill = group)) +
  geom_ribbon(
    aes(ymin = mean_k - se_k, ymax = mean_k + se_k),
    alpha = 0.15,
    color = NA
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_hline(yintercept = 16, linetype = "dotted", linewidth = 0.8) +
  scale_x_continuous(breaks = 1:10) +
  scale_color_manual(
    values = c(
      "Above optimum" = "darkorange3",
      "Below optimum" = "steelblue3"
    )
  ) +
  scale_fill_manual(
    values = c(
      "Above optimum" = "darkorange3",
      "Below optimum" = "steelblue3"
    )
  ) +
  labs(x = "Trial", y = "Search set size (k)", color = NULL, fill = NULL) +
  theme_minimal(base_size = 14)

# group × round
dat_group <- dat_analysis |>
  filter(!is.na(group)) |>
  mutate(
    round_c = round - mean(round),
    group_dummy = ifelse(group == "Above optimum", 0.5, -0.5)
  )

lmm_group <- lme(
  explore_count ~ round_c * group_dummy,
  random = ~ 1 | user_id,
  data = dat_group,
  method = "ML"
)
summary(lmm_group)
Anova(lmm_group, type = "3")

b_round <- fixef(lmm_group)["round_c"]
b_interaction <- fixef(lmm_group)["round_c:group_dummy"]

slope_above <- b_round + b_interaction * 0.5
slope_below <- b_round + b_interaction * -0.5
c(slope_above = slope_above, slope_below = slope_below)

## Robustness Check
dat_group_rtm <- dat_group |>
  filter(round > initial_round) |>
  mutate(round_c = round - mean(round))

lmm_group_rtm <- lme(
  explore_count ~ round_c * group_dummy,
  random = ~ 1 | user_id,
  data = dat_group_rtm,
  method = "ML"
)
summary(lmm_group_rtm)
Anova(lmm_group_rtm, type = "3")
```


# Testing for the optimal search set size in Lab

```{r}
k_mean <- mean(dat_analysis$explore_count, na.rm = TRUE)

dat_analysis <- dat_analysis |>
  mutate(explore_c = explore_count - k_mean)

lmm_reward <- lme(
  z_point ~ explore_c + I(explore_c^2),
  random = ~ 1 | user_id,
  data = dat_analysis,
  method = "REML",
  na.action = na.omit
)

b <- fixef(lmm_reward)
b1 <- b["explore_c"]
b2 <- b["I(explore_c^2)"]


k_emp_centered <- -b1 / (2 * b2)
k_emp <- k_emp_centered + k_mean
k_emp
```

```{r}
# Participant Cluster Bootstrap
set.seed(1)

boot_k <- replicate(2000, {
  ids <- sample(unique(dat_analysis$user_id), replace = TRUE)

  d <- purrr::map2_dfr(ids, seq_along(ids), function(uid, i) {
    dat_analysis |>
      dplyr::filter(user_id == uid) |>
      dplyr::mutate(boot_id = paste0(uid, "_", i))
  })

  fit <- try(
    lme(
      z_point ~ explore_c + I(explore_c^2),
      random = ~ 1 | boot_id,
      data = d,
      method = "REML"
    ),
    silent = TRUE
  )
  if (inherits(fit, "try-error")) {
    return(NA)
  }

  bb <- fixef(fit)
  b1_i <- bb["explore_c"]
  b2_i <- bb["I(explore_c^2)"]
  if (b2_i >= 0) {
    return(NA)
  }

  (-b1_i / (2 * b2_i)) + k_mean
})

quantile(boot_k, c(.025, .975), na.rm = TRUE)
```


```{r}
n_valid <- sum(!is.na(boot_k))
n_total <- length(boot_k)
n_valid
n_valid / n_total

hist(boot_k, breaks = 50)
summary(boot_k)
```

```{r}
summary(lmm_reward)
```

```{r}
lmm_linear <- lme(
  z_point ~ explore_c,
  random = ~ 1 | user_id,
  data = dat_analysis,
  method = "ML"
)
lmm_reward_ml <- update(lmm_reward, method = "ML")
AIC(lmm_linear, lmm_reward_ml)
```

```{r}
boot_valid <- boot_k[!is.na(boot_k)]

median(boot_valid)

k_range <- range(dat_analysis$explore_count, na.rm = TRUE)
boot_trunc <- boot_valid[boot_valid >= k_range[1] & boot_valid <= k_range[2]]

length(boot_trunc) / length(boot_valid)
quantile(boot_trunc, c(.025, .975))
median(boot_trunc) # reported median, same (truncated) distribution as the CI above
```


# SI:Within-Set Behavior
```{r}
dat_analysis |>
  mutate(
    trial = round - 3,
    session = ifelse(trial <= 5, "Early (Trial 1–5)", "Late (Trial 6–10)"),
    selected_best = theo_inner_regret == 0,
    k_bin = cut(
      explore_count,
      breaks = c(0, 10, 15, 20, 25, 30, Inf),
      labels = c("≤10", "11–15", "16–20", "21–25", "26–30", ">30")
    )
  ) |>
  group_by(k_bin, session) |>
  summarise(
    prob_best = mean(selected_best, na.rm = TRUE),
    n = n(),
    se = sqrt(prob_best * (1 - prob_best) / n),
    .groups = "drop"
  ) |>
  ggplot(aes(x = k_bin, y = prob_best, color = session, group = session)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = prob_best - se, ymax = prob_best + se),
    width = 0.2
  ) +
  scale_color_manual(
    values = c(
      "Early (Trial 1–5)" = "steelblue3",
      "Late (Trial 6–10)" = "darkorange3"
    )
  ) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  labs(
    x = "Search set size (k)",
    y = "P(selecting best within set)",
    color = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top")
```

```{r}
library(lme4)

### selected_best = observed_best == user_chosen_point
dat_ws <- dat_analysis |>
  mutate(
    selected_best = as.integer(theo_inner_regret == 0),
    k_z = as.numeric(scale(explore_count))
  )

mean(dat_ws$selected_best)

glmer_ws <- glmer(
  selected_best ~ z_round + k_z + (1 | user_id),
  data = dat_ws,
  family = binomial,
  na.action = na.omit
)
summary(glmer_ws)
```

```{r}
glm_ws <- glm(
  selected_best ~ z_round + k_z,
  data = dat_ws,
  family = binomial,
  na.action = na.omit
)
summary(glm_ws)

glm_ws_raw <- glm(
  selected_best ~ I(round - 3) + explore_count,
  data = dat_ws,
  family = binomial,
  na.action = na.omit
)
cbind(
  OR = exp(coef(glm_ws_raw)),
  exp(confint.default(glm_ws_raw))
)
```



# SI:Concentration of Within-Set Sampling

```{r}
# Participants × Trials × Options: View Count Rank and Percentage of Total Views (Out of 40)
vlong <- dat_analysis |>
  drop_na(observed_choice_order, round) |>
  separate_rows(observed_choice_order, sep = "_") |>
  filter(observed_choice_order != "") |>
  count(user_id, round, observed_choice_order, name = "view_count") |>
  group_by(user_id, round) |>
  arrange(desc(view_count), .by_group = TRUE) |>
  mutate(
    rank = row_number(),
    share = view_count / sum(view_count)
  ) |>
  ungroup() |>
  mutate(
    trial = round - 3,
    phase = ifelse(trial <= 5, "Early (trials 1–5)", "Late (trials 6–10)")
  )

vlong |>
  count(rank) |>
  mutate(
    frac = n /
      dplyr::n_distinct(paste(dat_analysis$user_id, dat_analysis$round))
  ) |>
  head(20)
```


### Changes in Sampling Within the search set for Each Trial
```{r}
library(tidyverse)

seq_len_check <- dat_analysis |>
  mutate(n_tokens = str_count(observed_choice_order, "_") + 1) |>
  count(n_tokens)
seq_len_check

# Entropy of the Distribution of sampling by User and Round
solid_data <- dat_analysis |>
  drop_na(observed_choice_order, round) |>
  separate_rows(observed_choice_order, sep = "_") |>
  filter(observed_choice_order != "") |>
  count(user_id, round, observed_choice_order, name = "view_count") |>
  group_by(user_id, round) |>
  mutate(p = view_count / sum(view_count)) |>
  summarise(
    entropy = -sum(p * log2(p)),
    n_options = dplyr::n(),
    total_views = sum(view_count),
    .groups = "drop"
  ) |>
  mutate(
    evenness = entropy / log2(n_options)
  )

# Point amd search set size
reward_data <- dat_analysis |>
  dplyr::select(user_id, round, user_chosen_point, explore_count) |>
  distinct()

analysis_df <- solid_data |>
  left_join(reward_data, by = c("user_id", "round")) |>
  mutate(
    z_point = as.numeric(scale(user_chosen_point)),
    z_entropy = as.numeric(scale(entropy)),
    z_evenness = as.numeric(scale(evenness)),
    z_k = as.numeric(scale(explore_count)),
    z_k2 = as.numeric(scale(explore_count^2)),
    z_round = as.numeric(scale(round))
  )

cor(analysis_df[c("entropy", "evenness", "explore_count")])
```

### SI figure 1: allocation shape by trial 
```{r}
library(ggplot2)
library(ggeffects)

pal <- "#c0392b"
MAXR <- 15

fd <- vlong |>
  filter(rank <= MAXR) |>
  group_by(trial, rank) |>
  summarise(m = mean(share), se = sd(share) / sqrt(n()), .groups = "drop") |>
  mutate(trial_f = factor(paste("Trial", trial), levels = paste("Trial", 1:10)))
ref <- fd |>
  filter(trial == 1) |>
  dplyr::select(rank, ref = m) |>
  tidyr::crossing(
    trial_f = factor(paste("Trial", 1:10), levels = paste("Trial", 1:10))
  )

fig_allocation <- ggplot(fd, aes(rank, m)) +
  geom_col(aes(fill = m), width = 0.9) +
  geom_line(
    data = ref,
    aes(rank, ref),
    colour = "grey35",
    linewidth = 0.5,
    linetype = "dashed"
  ) +
  facet_wrap(~trial_f, nrow = 2) +
  scale_fill_viridis_c(option = "mako", direction = -1, guide = "none") +
  scale_x_continuous(breaks = c(1, 5, 10, 15)) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Option rank within a trial  (1 = most sampled)",
    y = "Mean share of the 40 views",
    tag = "a"
  ) +
  theme_classic(base_size = 14) + # Figure 1 と同じ
  theme(
    plot.tag = element_text(face = "bold", size = 17),
    panel.spacing = unit(0.7, "lines"),
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 13)
  )
fig_allocation
```

### SI figure 2: concentration → payoff 
```{r}
m_pay <- lmer(
  z_point ~ z_evenness + z_k + z_k2 + z_round + (1 | user_id),
  data = analysis_df
)

pay_m <- mean(analysis_df$user_chosen_point)
pay_sd <- sd(analysis_df$user_chosen_point)
sd_e <- sd(analysis_df$evenness)
m_e <- mean(analysis_df$evenness)
q <- as.numeric(quantile(analysis_df$evenness, c(.05, .95)))
zs <- (seq(q[1], q[2], length = 80) - m_e) / sd_e
prd <- as.data.frame(
  ggpredict(
    m_pay,
    terms = sprintf("z_evenness [%s]", paste(zs, collapse = ","))
  )
) |>
  mutate(
    evenness = x * sd_e + m_e,
    across(c(predicted, conf.low, conf.high), ~ .x * pay_sd + pay_m) # z → 素点
  )

fig_concentration_payoff <- ggplot(prd, aes(evenness, predicted)) +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high),
    fill = pal,
    alpha = 0.18
  ) +
  geom_line(colour = pal, linewidth = 1.2) +
  scale_x_reverse(labels = scales::number_format(accuracy = 0.01)) +
  labs(
    x = expression(
      "Evenness of allocation  " *
        (italic(H) / log[2] * italic(k)) *
        "   (" *
        symbol('\254') *
        " more concentrated)"
    ),
    y = "Model-predicted payoff (points)",
    tag = "b"
  ) +
  theme_classic(base_size = 14) + # Figure 1 と同じ
  theme(plot.tag = element_text(face = "bold", size = 17))
fig_concentration_payoff
```

```{r}
trend_ev <- summary(lmer(
  z_evenness ~ z_round + (1 | user_id),
  data = analysis_df
))$coefficients["z_round", ]
trend_en <- summary(lmer(
  z_entropy ~ z_round + (1 | user_id),
  data = analysis_df
))$coefficients["z_round", ]
pay_ev <- summary(m_pay)$coefficients["z_evenness", ]
pay_en <- summary(
  lmer(
    z_point ~ z_entropy + z_k + z_k2 + z_round + (1 | user_id),
    data = analysis_df
  )
)$coefficients["z_entropy", ]
cap <- function(tag, co) {
  cat(sprintf(
    "%-30s β = %+.2f, SE = %.2f, t(%.0f) = %+.2f, p = %.3g\n",
    tag,
    co[1],
    co[2],
    co[3],
    co[4],
    co[5]
  ))
}
cap("Fig 1  evenness ~ trial", trend_ev)
cap("       entropy  ~ trial", trend_en)
cap("Fig 2  payoff ~ evenness", pay_ev)
cap("       payoff ~ entropy", pay_en)
```

```{r}
ggplot(analysis_df, aes(explore_count, entropy)) +
  geom_point(alpha = 0.2, size = 0.9, colour = "grey35") +
  geom_smooth(
    method = "loess",
    formula = y ~ x,
    colour = "grey20",
    se = FALSE,
    linewidth = 0.8
  ) +
  labs(
    x = "Search-set size (k)",
    y = "Raw entropy (bits)",
    subtitle = paste0(
      "r(entropy, k) = ",
      round(cor(analysis_df$entropy, analysis_df$explore_count), 2),
      "   |   r(evenness, k) = ",
      round(cor(analysis_df$evenness, analysis_df$explore_count), 2)
    )
  ) +
  theme_classic(base_size = 12)
```


### LMM
```{r}
library(lme4)
library(lmerTest)

model_trend <- lmer(z_entropy ~ z_round + (1 | user_id), data = analysis_df)
summary(model_trend) # β for z_round

model_trend_even <- lmer(
  z_evenness ~ z_round + (1 | user_id),
  data = analysis_df
)
summary(model_trend_even)

lmer(entropy ~ I(round - 3) + (1 | user_id), data = analysis_df) |>
  fixef() # b = bits per trial
```

```{r}
library(car)

summary(lmer(z_point ~ z_entropy + z_round + (1 | user_id), data = analysis_df))

model_reward_k2 <- lmer(
  z_point ~ z_entropy + z_k + z_k2 + z_round + (1 | user_id),
  data = analysis_df
)
summary(model_reward_k2)
vif(model_reward_k2)

model_reward_even <- lmer(
  z_point ~ z_evenness + z_k + z_k2 + z_round + (1 | user_id),
  data = analysis_df
)
summary(model_reward_even)
vif(model_reward_even)
```
